### 2015-07-02 Manuscript/supplementary.tex
Updates to supp plots.

\documentclass{article} 
\usepackage{graphicx}

\title{Supplementary figures}

\begin{document}

\maketitle

Supplemental figures 1 and 2 are analogous to figures 1 and 2 from the main text, but with habitat patches arranged in a lattice structure (i.e. each habitat patch is accessible only from its four nearest neighbors).

Supplemental figures 3 and 4 show within-patch occupancy for a given $\nu$ and $\delta$.

\begin{figure}
\includegraphics[scale=0.5]{figure/supplement_1.pdf}
\caption{The proportion of simulations resulting in endemic disease dynamics (persistence of both susceptible and infectious hosts) as a function of disease-induced mortality ($\nu$), and the probability of direct transmission ($\delta$).  Panel columns show low (0.3, a, d), medium (1.0, b, e), and high (3.2, c, f) pathogen longevities, while rows show high quality (ranging from 0.76 to 1.77, a - c) and low quality (ranging from 0.23 to 1.24, d - f) patch quality distributions.  Darker shading corresponds to more frequent endemic dynamics.  The white region above the endemic region generally corresponds to pandemic dynamics (no susceptible host populations persist), while the region below the endemic region corresponds to disease-free dynamics.}
\label{endemic}
\end{figure}

\begin{figure}
\includegraphics[scale=0.5]{figure/supplement_2.pdf}
\centering
\caption{Differences between the median total effective population size of hosts on high quality patch distributions (quality between 0.76 and 1.77) and low quality patch distributions (quality between 0.23 to 1.24) as a function of disease-induced mortality ($\nu$) and the probability of direct transmission ($\delta$).  Reds indicate higher median host population sizes in high quality patch distributions, while blues indicate higher median host population sizes in low quality patch distributions.  Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.}
\label{highvlow}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_3}
\centering
\caption{Probability of susceptible occupancy (measured as the proportion of time spent occupied by susceptible hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Disease-induced mortality and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_4}
\centering
\caption{Probability of infectious occupancy (measured as the proportion of time spent occupied by infectious hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Disease-induced mortality and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}

\end{document}